\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amscd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{enumerate}
\usepackage[all,cmtip]{xy}

\textwidth=15cm \textheight=22cm \topmargin=0.5cm \oddsidemargin=0.5cm \evensidemargin=0.5cm

\newcommand{\sk}{\vskip 10mm}
\newcommand{\bb}[1]{\mathbb{#1}}

\theoremstyle{plain}
\newtheorem{problem}{Problem}
\newtheorem{lemma}{Lemma}[problem]

\theoremstyle{remark}
\newtheorem{tpart}{}[problem]
\newtheorem*{ppart}{}

\begin{document}

\begin{problem}[12]
  Show that the quotient map $S^1\times S^1\rightarrow S^2$ collapsing the
  subspace $S^1\vee S^1$ to a point is not nullhomotopic
  by showing that it induces an isomorphism on $H_2$. On the
  other hand, show via covering spaces that any map
  $S^2\rightarrow S^1\times S^1$ is nullhomotopic.
\end{problem}

\begin{proof}
  
\end{proof}

\sk

\begin{problem}[14]
  A map $f:S^n\rightarrow S^n$ satisfying $f(x)=f(-x)$ for all $x$ is called
  an even map. Show that an even map $S^n\rightarrow S^n$ must have even
  degree, and that the degree must in fact be zero when $n$ is even.
  When $n$ is odd, show that there exist even maps of any given even
  degree. [Hints: If $f$ is even, it factors as a composition
  $S^n\rightarrow\bb{R}P^n\rightarrow S^n$. Using the calculation of
  $H_n(\bb{R}P^n)$ in the text, show that the induced map
  $H_n(S^n)\rightarrow H_n(\bb{R}P^n)$ sends a generator to twice a
  generator when $n$ is odd. It may be helpful to show that the
  quotient map $\bb{R}P^n\rightarrow \bb{R}P^n/\bb{R}P^{n-1}$ induces an
  isomorphism on $H_n$ when $n$ is odd.]
\end{problem}

\begin{proof}
    Let $f$ be a map from $S^n$ to $S^n$ such that $f(x)=f(-x)$ for all $x\in S^n$.
  Since $\bb{R}P^n$ is a quotient space of $S^n$ where antipodal points are
  identified any even map from $S^n$ respects equivalence classes for the
  quotient space and as such it factors as
  \[
    \xymatrix{
      S^n \ar[d]^q \ar[r]^f & S^n\\
      \bb{R}P^n \ar[ru]_{\widetilde{f}}
    }
  \]

  First note that $\deg q =2$. This is because
  for any point $x\in\bb{R}P^n$ it will have two points mapping to it from $S^n$ and
  it will be the identity map giving us local degrees of 1 which add up to 2. Since
  any even map will factor in this way any even map must have even degree.
  
  When $n$ is even $H_n(\bb{R}P^n)\cong\bb{Z}$ and when $n$ is even $H_n(\bb{R}P^n)\cong 0$.
  If $n$ is odd and $f$ is even then $\deg f = \deg q\cdot\deg \widetilde{f}$. However
  since $H_n(\bb{R}P^n)\cong 0$ when $n$ is even then the degree of $f$ has to be zero
  as $\deg \widetilde{f}=0$ since it is mapping out of the trivial group.

  Now suppose that $n$ is odd. There is a quotient map
  $r:\bb{R}P^n\rightarrow (\bb{R}P^n/\bb{R}P^{n-1}\equiv S^n)$. This map will have degree $1$ and
  as such $q\circ r$ will be a map from $S^n\rightarrow S^n$ of degree two. From there if we
  take a map of degree $k$, $f_k:S^n\rightarrow S^n$ (Hatcher $2.32$). Then $f_k\circ q\circ r$ will be
  an even map of degree $2k$.
\end{proof}

\sk

\begin{problem}[20]
  For finite CW complexes $X$ and $Y$, show that
  $\chi(X\times Y)=\chi(X)\chi(Y)$.
\end{problem}

\begin{proof}
  Let $x_n$ and $y_n$ denote the $n$-dimensional simplices of $X$ and $Y$
  respectively. Then
  \[
    \chi(X)\chi(Y)=\left(\sum_{n=0}^\infty(-1)^ix_i\right)\left(\sum_{m=0}^\infty(-1)^n\right)=\sum_{0=i=m+n}^\infty(-1)^{i}\left(\sum_{j=0}^ix_jy_{i-j}\right)
  \]

  However since the number of $i$-simplices in $X\times Y$ is $\sum_{j=0}^ix_jy_{i-j}$
  this demonstrates that
  \[
    \chi(X)\chi(Y)=\chi(X\times Y)
  \]
\end{proof}

\sk

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
