\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amscd}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{enumerate}
\usepackage[all,cmtip]{xy}

\textwidth=15cm \textheight=22cm \topmargin=0.5cm \oddsidemargin=0.5cm \evensidemargin=0.5cm

\newcommand{\sk}{\vskip 10mm}
\newcommand{\bb}[1]{\mathbb{#1}}
\newcommand{\ra}{\rightarrow}
\newcommand{\rh}{\widetilde{h}}

\theoremstyle{plain}
\newtheorem{problem}{Problem}
\newtheorem{lemma}{Lemma}[problem]

\theoremstyle{remark}
\newtheorem{tpart}{}[problem]
\newtheorem*{ppart}{}

\begin{document}

\begin{problem}[2.2.36]
  Show that $H_i(X\times S^n)\cong H_i(X)\oplus H_{i-n}(X)$ for all $i$ and $n$,
  where $H_i=0$ for $i<0$ be definition. Namely, show
  $H_i(X\times S^n)\cong H_i(X)\oplus H_i(X\times S^n,X\times\{x_0\})$ and
  $H_i(X\times S^n,X\times\{x_0\})\cong H_{i-1}(X\times S^{n-1},X\times\{x_0\})$
  [For the latter isomorphism the relative Mayer-Vietoris sequence yields
  an easy proof].
\end{problem}

\begin{proof}
  First we show that $H_i(X\times S^n)\cong H_i(X)\oplus H_i(X\times S^n,X\times\{ x_0\})$. Swap out
  $H_i(X)$ for $H_i(X\times\{x_0\})$. Then apply the long exact sequence for the pair to get
  \[
    \cdots\rightarrow H_{i+1}(X\times S^n,X\times\{ x_0\}) \rightarrow H_i(X\times\{x_0\}) \rightarrow H_i(X\times S^n) \rightarrow H_i(X\times S^n,X\times\{ x_0\}) \rightarrow H_{i-1}(X\times\{x_0\}) \rightarrow \cdots
  \]
  This splits into short exact sequences of the form
  \[
    0\rightarrow H_i(X)\rightarrow H_i(X\times S^n)\rightarrow H_i(X\times S^n,X\times \{x_0\})\rightarrow 0
  \]
  Since $X\times\{x_0\}$ is a retract of $X\times S^n$ we have that the sequence splits.
  Thus by the splitting lemma we have that $H_i(X\times S^n)\cong H_i(X)\oplus H_i(X\times S^n,X\times\{ x_0\})$.

  For the next isomorphism $H_i(X\times S^n,X\times\{x_0\})\cong H_{i-1}(X\times S^{n-1},X\times\{x_0\})$ apply
  the relative Mayer-Vietoris sequence with $A=X\times S^n\setminus\{x_0\}$, $B=X\times S^n\setminus\{x_1\}$ ($x_1\neq x_0$),
  and $C,D=X\times \{x_0\}$. Then we have
  \[
    \cdots\rightarrow H_i(X\times S^n\setminus\{x_0\},X\times\{x_0\})\oplus H_i(X\times S^n\setminus\{x_0\},X\times\{x_1\}) \rightarrow H_i(X\times S^n,X\times\{x_0\}) 
  \]
  \[
    \rightarrow H_{i-1}(X\times S^{n-1},X\times\{x_0\})\rightarrow H_{i-1}(X\times S^n\setminus\{x_0\},X\times\{x_0\})\oplus H_{i-1}(X\times S^n\setminus\{x_0\},X\times\{x_1\}) \rightarrow\cdots
  \]

  However the terms on the left and the right are all zero. As such
  this induces an isomorphism between $H_i(X\times S^n,X\times\{x_0\})\cong H_{i-1}(X\times S^{n-1},X\times\{x_0\})$.

  Finally apply the second isomorphism $n$ times to the first isomorphism to get
  $H_i(X\times S^n)\cong H_i(X\times\{x_0\})\oplus H_{i-n}(X\times S^0,X\times\{x_0\})$. However the relative homology
  will make one of the two copies of $X$ in $X\times S^0$ null. As such
  $H_{i-n}(X\times S^0,X\times\{x_0\})\cong H_{i-n}(X)$.

  Therefore $H_i(X\times S^n)\cong H_i(X)\oplus H_{i-n}(X)$ for all $i,n$ where we assume $H_i(X)=0$ for
  $i<0$.
\end{proof}

\sk

\begin{problem}[2.3.3]
  Show that if $\rh$ is a reduced homology theory then $\rh_n(point)=0$ for all
  $n$. Deduce that there are suspension isomorphisms $\rh_n(X)\cong\rh_{n+1}(SX)$
  for all $n$.
\end{problem}

\begin{proof}
  First note that if we wedge the point with itself that we get a point back.
  Then using the wedge axiom we have $\rh_i(\bigvee_1^2*=*)$ is isomorphic to
  $\rh_i(*)\oplus\rh_i(*)$. The only way this could occur would be if $\rh_i(*)\cong 0$
  for all $i$.

  For the suspension use the Mayer-Vietoris sequence along with the
  fact that the cone of a space is contractible to get
  \[
    \xymatrix{
      \cdots \ar[r] & \rh_{i+1}(CX)\oplus \rh_{i+1}(CX) \ar[r] & \rh_{i+1}(SX) \ar[r] & \rh_i(X) \ar[r] & \rh_i(CX)\oplus \rh_i(CX) \ar[r] & \cdots
    }
  \]
  Which is equivalent to
  \[
    \xymatrix{
      0 \ar[r] & \rh_{i+1}(SX) \ar[r] & \rh_i(X) \ar[r] & 0
      }
  \]
  This show that $\rh_i(X)\cong\rh_{i+1}(SX)$ for all $i$.
\end{proof}

\sk

\begin{problem}[2.B.3]
  Let $(D,S)\subset (D^n,S^{n-1})$ be a pair of subspaces homeomorphic to
  $(D^k,S^{k-1})$, with $D\cap S^{n-1}=S$. Show that the inclusion
  $S^{n-1}-S\hookrightarrow D^n-D$ induces an isomorphism on homology. [Glue two
  copies of $(D^n,D)$ to the two ends of $(S^{n-1}\times I,S\times I)$ to
  produce a $k$-sphere in $S^n$ and look at the Mayer-Vietoris sequence for the
  complement of this $k$-sphere.]
\end{problem}

\begin{proof}
  Using the construction listed in the problem decompose $S^n\setminus S^k$ as $A\cup B$
  where $A=(S^{n-1}\times I,S\times I)$ filled in with $D^n\setminus D$ on one side and $B$ as the
  filling on the other. Then using the fact that taking the Cartesian product
  with the interval does not change the homology and applying the Mayer-Vietoris
  sequence we get:
  \[
    \xymatrix{
      \cdots \ar[r] & H_i(S^{n-1}\setminus S) \ar[r] & H_i(D^n\setminus D)\oplus H_i(D^n\setminus D) \ar[r] & H_i(S^n\setminus S^k) \ar[r] & \cdots
    }
  \]

  However $H_i(S^n\setminus S^k)=0$ unless $i=n-k-1$. Likewise since $S\cong S^{k-1}$ as a
  subspace we have that $H_i(S^{n-1}\setminus S)\cong 0$ unless $i=(n-1)-(k-1)-1=n-k-1$.
  Similarly $H_i(D^n\setminus D)=0$ unless $i=n-k-1$. They will all be $\bb{Z}$ otherwise.
  Thus we get the short exact sequence
  \[
    \xymatrix{
      \cdots \ar[r] & H_{n-k-1}(S^{n-1}\setminus S) \ar[r]^-{i_*,i_*} & H_{n-k-1}(D^n\setminus D)\oplus H_{n-k-1}(D^n\setminus D) \ar[r]^-{k_*-k_*} & H_{n-k-1}(S^n\setminus S^k) \ar[r] & \cdots
    }
  \]

  From this we can see that the kernel of $k_*-k_*$ is $\{(x,x)\in \bb{Z}^2\}$.
  However since $(1,1)\in\ker (k_*-k_*)$ from exactness it must be that
  $i_*(1)=1$ which shows that $i_*$ is an isomorphism. Since $i_*$ is the
  zero map between groups that are zero elsewhere this shows that the
  map induced by inclusion is indeed an isomorphism for all $i$.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}
